>

### 2019-10-02 manuscript/manuscript.Rmd
Revise methodology section

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers can inhibit holistic analysis,
may have been inititially motivated by computational limitations [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of a point is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_ij)$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function. 
@Giles-Corti2005 compare gravity-based accessibility scores with an isochrone
specification and show that the former are more predictive of walking behavior;
that is, the attributes of a park matter influence walking more than merely the
distance to it. @Zhang2011 present a population-weighted gravity-based
accessibility to parks metric with a national rather than metropolitan scope, 
though they do not examine its correlation with health outcomes.

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities* $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utility}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
the census tract $i$ in miles. The coefficients $\vect{\beta}$ can be
are frequently estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

@@

Both gravity and utility-based specifications hold several advantages relative
to isochrone-based accessibility metrics more commonly found in the literature.
First, all individuals are defined as having some access to all parks, rather
than an arbitrary cutoff asserted by the researcher. This allows for the fact
that some people are more or less sensitive to distances, and that distance is a
continuous, non-binary phenomenon. It defies reason to assume a person living
1.1 miles from a park has functionally different accessibility than someone
living at 0.9 miles. Second, the random utility formulation allows the
researcher to include any attribute of the park as part of its utility
specification. This suggests that not all parks are equal, and that a large park
such as Central Park in Manhattan may provide health and activity benefits over
a much larger area than a smaller community square.

@@

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We also consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these sub-facilities, we created 
variables for each park indicating the presence of sports courts, playgrounds,
and trails.

@@

@@

The spatial weights matrix $W$ is constructed of individual elements where
$w_{ij} = 0$ if observations $i$ and $j$ are spatially independent of each other,
and $w_{ij} > 0$ if $i$ and $j$ are spatially related in some way. 
@Dubin1998 presents details on constructing $W$, but in this study we
assume that tracts sharing at least one common border point are neighbors of
each other. The elements of $W$ are then row-standardized so that each observation
has an equal total influence from all its neighbors.

@@

Extant destination choice models used in practice typically use travel time in
minutes as a travel impedance term, and employment by sector as a size term or
attraction component. Typically such models handle recreational trips together with
other non-work and non-school trips, and we therefore can find no previously
estimated coefficients using size terms relevant to park acreage or amenities.
Given this gap in the literature, we instead assert plausible values for
$\beta$ to generate the choice-based accessibility statistic used in this study.

The functional form of the accessibility statistic is adapted from
Equation \ref{eq:v_sd} by log-transforming first the distance (Euclidean, in miles)
between the park boundaries and population-weighted tract centroids and second the
park size (in acres).
\begin{equation}\label{eq:v_logged}
V_{ij} = \log(size_j)\lambda_s + \log(distance_{ij})\lambda_d
\end{equation}
We then calculate the consumer surplus $C$  (log-sum of the exponentiated
utilities) for each tract and standardize the result,
\begin{equation}\label{eq:v_logged}
C_i' = \frac{C_i - \bar{\vect{C}}}{sd(\vect{C})}
\end{equation}

We initially estimated potential values of the $\lambda$ coefficients by
applying a limited-memory, box-constrained optimization algorithm
maximizing the model log-likehood.\footnote{Using the {\texttt nlm} function
in R [@RCoreTeam2018]} The constraint required that the coefficient on
size be positive and that on distance negative; all else equal, we assert people
will prefer to use larger parks and parks that are nearer to their residence. This
resulted in $\lambda_d = -1.98$  and $\lambda_s = 0.000$. Though the ``most
likely'' parameter value, in this case it was unsatisfactory for two reasons:
first, our prior assertion that park size is a valuable attribute of a park, and
second the observation that with these maximum likelihood values tracts
bordering on the corners of Central Park were measured with poor overall
accessibility, an unacceptably unintuitive result. We subsequently adjusted the
$\lambda_s$ value to $0.3$; the relative ratio of the coefficients implies that
all else equal, an individual would travel 6.6 times further to access a park
twice as large.

Figure \ref{fig:map} shows a map of New York City with each tract highlighted
based on its relative accessibility score. The most continuous region of good park
access is in upper Manhattan and the Bronx, bracketed by Central Park and the
Bronx River. Conversely, some of the poorest-accessibility areas are in
Brooklyn tracts not immediately adjacent to Prospect Park. Because the
accessibility statistic is normalized, the worst values are slightly below $-3$,
the best somewhere above $3$.

>>>>

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of a point is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_ij)$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function. 
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park matter influence walking more than
merely the distance to it. @Zhang2011 present a population-weighted
gravity-based accessibility to parks metric with a national rather than
metropolitan scope, though they do not examine its correlation with health
outcomes.

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities* $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
The coefficients $\vect{\beta}$ are frequently estimated from household
surveys, though in the absence of a survey we may assert reasonable values.

@@

Both gravity and utility-based specifications hold several advantages relative
to isochrone-based accessibility metrics more commonly found in the literature.
First, all individuals are defined as having some access to all parks, rather
than an arbitrary cutoff asserted by the researcher. This allows for the fact
that some people are more or less sensitive to distances, and that distance is a
continuous, non-binary phenomenon. It defies reason to assume a person living
1.1 miles from a park has functionally different accessibility than someone
living at 0.9 miles. Second, the random utility formulation allows the
researcher to include -- in principle -- any attribute of the park as part of
its utility specification. This suggests that not all parks are equal, and that
a large park with many amenities such as Central Park in Manhattan may provide
health and activity benefits over a much larger area than a smaller community
square.

@@

In a destination choice framework, the tracts represent the "origins" and the
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We also consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these distinct sub-facilities, we created 
variables for each containing park indicating the presence of sports courts,
playgrounds, and trails.

@@

We calculate a set of utility-based accessibility statistics for each tract
in New York City. The most basic utility specification includes only the
the park size in acres and the distance from the population-weighted tract centroid
to the boundary of the park in miles,
\begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation}
The logarithmic transform allows for diminishing marginal utility of distance
and park size: A 1-mile increase to a trip matters more for a 1-mile trip than a
10-mile trip. @Macfarlane2019 estimated destination choice 
parameters for park trips in the Alameda County (Oakland), California using
this utility specification, obtaining values of $\lambda_s = 0.373$ and 
$\lambda_d = -1.98$. The ratio of these estimates implies people are willing to
travel roughly six times further to reach a park twice as large. In the absence 
of park trip destination choice coefficients in New York City, we apply these
previous estimates.

The second utility specification is adapted from a park destination choice model
estimated by @Kinnell2006 in New Jersey. In this study the authors used a revealed
preference survey to determine which factors of a park and its surrounding environment
affected its selection. By bringing over common variables, we can create a second
utility specification as
\begin{equation}\label{eq:u_multi}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_t * trails + \\
  \lambda_{c} * courts + \lambda_p * playgrounds
\end{equation}
with $\lambda_t = 0.99$, $\lambda_c = 0.43$, and $\lambda_p = 0.26$. @Kinnell 
did not transform their size and distance estimates, so we retain the previously 
estimated and applied values. We assume that the other covariates in the @Kinnell
model which we do not have available to us (e.g., boat launches, picnic areas) are
orthogonal to the included parameters and leaving them out will not affect the
values of the included coefficients.

The final utility specification includes the number of geolocated tweets
emanating from within the park in September of 2014 in addition to the
basic specification in Equation \ref{eq:u_access},
\begin{equation}\label{eq:u_tweets}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_{tw} * YJ(tweets)
\end{equation}
with $YJ(x)$ being a Yeo-Johnson transformation that implements a diminishing
marginal return similar to $log(x)$, but where $YJ(0) = 0$ [@Yeo2000]. In this
case we have no external information describing how destination choice to
parks may be affected by the twitter activity in the park, so we assert a 
value of $\lambda_{tw} = 0.1$

We then calculate the utility-based accessibility of each tract $A_i$ as defined
in Equation \ref{eq:utility} under each utility specification. Recall that the 
total value of the accessibility is relative to an unknown scalar $C$; for this 
reason we standardize the accessibility values for all tracts within each
utility specification, 
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{A}}{sd(A)}
\end{equation}

Figure \ref{fig:map} shows a map of New York City with each tract highlighted
based on its relative basic utility-based accessibility score (Equation
\ref{eq:u_access}). The most continuous region of good park access is in upper
Manhattan and the Bronx, bracketed by Central Park and the Bronx River.
Conversely, some of the poorest-accessibility areas are in Brooklyn tracts not
immediately adjacent to Prospect Park. Because the accessibility statistic is
normalized, the worst values are slightly below $-3$, the best somewhere above
$3$.

For comparative purposes, we also employ an isochrone analysis using the
10-minute walk threshold calculated for ParkScore [@parkscore2019]. If the
population-weighted centroid of a tract is within a 10-minute walk of a green space
as defined by ParkScore, the tract is considered as having "access" to a park.
The centroid is necessary as all tracts in New York City have at least *some* 
intersection with the 10-minute walk buffer; as it is, only 
`r sum(!tracts$walk_10)` of the `r nrow(tracts)` tract centroids are not located 
in this buffer.

@@

The spatial weights matrix $W$ is constructed of individual elements where
$w_{ij} = 0$ if observations $i$ and $j$ are spatially independent of each other,
and $w_{ij} > 0$ if $i$ and $j$ are spatially related in some way. 
@Dubin1998 presents details on constructing $W$, but in this study we
assume that tracts sharing at least one common border point are neighbors of
each other^[We also considered a distance-discounted weights matrix, which
produced similar results. This matrix is selected for simplicity.]. The elements
of $W$ are then row-standardized so that each observation has an equal total
influence from all its neighbors.

@@
